# Reproducible Research: Peer Assessment 1
## Read in the data file and preprocess
```{r, echo=TRUE}
ad <- read.csv("activity.csv")
```
Variables are easier to manipulate if they are transformed into a slightly more user friendly format.
```{r, echo=TRUE}
steps <- ad$steps
date <- ad$date
interval <- ad$interval
```
You may want to get a visual idea of the distribution of steps taken in a day, so building a histogram will be helpful in examining these data.
## Building a Histogram
```{r, echo=TRUE}
hist(tapply(steps, date, sum), xlab="Steps", ylab= "Days", breaks=10, main="Steps taken on a Daily basis")
```
## Calculate the Mean and Median Values
Examination of the mean and median can be helpful in getting a good handle on the measures of central tendency in the data.We first create a framework that allows us to calculate the mean or median per day:
```{r, echo=TRUE}
steps_day <- as.numeric(tapply(steps, date, sum))
mean_steps_day <- mean(steps_day, na.rm=TRUE)
median_steps_day <- median(steps_day, na.rm=TRUE)
```
The median of the total number of steps taken per day is 1.0765 × 104Thus, the median and the mean values are very close together.
## Compute the average daily activity pattern
We want to build a time series plot at five minute intervals in order to obesrve the activity pattern.  You'll first want to change the interval variable to a factor variable and calculate the mean - as we have done above.
```{r, echo=TRUE}
interval <- as.factor(as.character(interval))
mean_int <- as.numeric(tapply(steps, interval, mean, na.rm=TRUE))
trueint <- data.frame(trueint = as.numeric(levels(interval)),mean_int)
trueint <- trueint[order(trueint$trueint),]
```
## Build the plot
```{r, echo=TRUE}
plot(trueint$trueint, trueint$mean_int, type = "l", main = "Average steps per minute", xlab="Time of Day", ylab="Mean Steps")
```
## Build the plot
```{r, echo=TRUE}
aveintper <- trueint[order(trueint$mean_int, decreasing=TRUE),]
head(aveintper, n=1)
```
## Input Missing Values
```{r, echo=TRUE}
missvals <- sum(is.na(steps))
steps[is.na(steps)] <- 37.38
summary(steps)
hist(tapply(newad$steps, newad$date, sum), xlab="Steps", ylab= "Days", breaks=10, main="Steps taken on a Daily basis")
```
```{r, echo=TRUE}
newsteps_day <- as.numeric(tapply(newad$steps, newad$date, sum))
newmean_steps_day <- mean(newsteps_day, na.rm=TRUE)
newmedian_steps_day <- median(newsteps_day, na.rm=TRUE)
newad_new <- newad
newad_new$date <- as.Date(newad_new$date)
newad_new$day <- weekdays(newad_new$date) == "Sunday"  | weekdays(newad_new$date) == "Saturday"
```
```{r, echo=TRUE}
newad_new_weekday <- newad_new[newad_new$day == FALSE, ]
newad_new_weekend <- newad_new[newad_new$day == TRUE, ]
#First create weekday and weekend variables
mean_weekday <- tapply(newad_new_weekday$steps, newad_new_weekday$interval, mean)
mean_weekend <- tapply(newad_new_weekend$steps, newad_new_weekend$interval, mean)
#then create two separate plots
```
```{r, echo=TRUE}
plot(mean_weekday, type = "l", xlab="Time interval", ylab="Steps", main="Weekday activity", col="darkblue")
plot(mean_weekend, type="l", xlab="Time interval", ylab="Steps", main="Weekend activity", col="red")
```
